### 2013-06-11 doc/dissertation/document/chapters/introduction/introduction.tex
adding some example figures for domain decomposed monte carlo

>>>>

\subsection{Solutions for the $SP_N$ Equations}
\label{subsec:spn_motiviation}
The neutron transport problem is complicated. Solutions cover a large
phase space and the problems of interest are often geometrically
complex, very large, or both, requiring tremendous computational
resources to generate an adequate solution. Modern deterministic
methods for large scale problems are commonly variants on the discrete
ordinates ($S_N$) method \citep{evans_denovo:_2010}. For fission
reactor neutronics simulations, the $S_N$ method requires potentially
trillions of unknown angular flux moments to be computed to achieve
good accuracy for the responses of interest
\citep{slaybaugh_acceleration_2011}. Other forms of the transport
problem, including the $P_N$ method, take on a simpler form than the
more common $S_N$ methods but lack in accuracy when compared while
still requiring considerable computational resources for solutions in
multiple dimensions.

In the 1960's, Gelbard developed an ad-hoc multidimensional extension
of the simple single dimension planar $P_N$ equations that created a
system of coupled, diffusion-like equations known as the simplified
$P_N$ ($SP_N$) equations. Up until around the 1990's, the $SP_N$
method was either widely unknown, widely unused, or combination of
both even though numerical studies showed promising results with
better solutions than diffusion theory and a significant reduction in
computational time over more accurate methods such as discrete
ordinates. Why did this happen? A significant problem, pointed out by
Brantley and Larsen \citep{brantley_simplified_2000}, was that little
rigor had been applied to the formulation of the $SP_N$ equations
since their derivation through primarily heuristic arguments. Instead,
studies at that time focused on simply comparing the results of the
method to other contemporary transport solution strategies. In
addition, many problems of interest from the literature at the time
were either solved using nodal-type methods for reactor-sized problems
or $S_N$-type methods for benchmark problems with intricate material
configurations and potentially large flux gradients over small spatial
domains.

So why reconsider the $SP_N$ equations? Starting in the 1990's and
primarily due to Larsen and his colleagues, the $SP_N$ equations have
been given a more rigorous treatment with both variational and
asymptotic derivations performed as a means of verification. In
addition, these equations have been more rigorously studied as
solution methods to MOX fuel problems and have been shown to provide
accurate solutions. With this mathematical literature to provide a
solid numerical footing for the method, we look at its application to
today's challenge problems in neutron transport for fission reactor
analysis. The reduction in numerical complexity of current
deterministic solution methods using the $S_N$ approximation could
mean significant savings in both compute time and memory required. In
addition, the characteristics of the solution to the transport problem
for a steady state reactor core permit diffusion theory to be used; a
staple of the nuclear industry since its inception. Therefore, if
diffusion theory is applicable, then finer grained solutions that
capture more of the physics contained in the transport equation should
be possible with the $SP_N$ method. In doing so, we also expect from
the literature to obtain computed responses on the order of accuracy
we would expect from an appropriately discretized $S_N$ method at a
fraction of the cost.

In order to leverage MCSA as a solution technique for physics
problems, its current formulation requires the linear operator and all
preconditioners to be explicitly formed. Recent developments in the
Exnihilo neutronics package at Oak Ridge National Laboratory have
permitted generation of the $SP_N$ system of equations for detailed
neutronics models of fission reactor-based
systems\citep{evans_simpli_2013}. By fully forming these equations and
formulating them as a linear algebra problem instead of using the
explicit iterative methods of the past, we now have access to all of
the modern advancements in computational linear algebra including
Krylov solvers for asymmetric systems and algebraic preconditioning
methods. This leads us to then explore the applicability of our work
in discrete Monte Carlo methods for linear systems as a possible
solution method for the $SP_N$ equations. In addition, solving the
$SP_N$ equations in this way also breaks away from the $S_N$ forms of
parallelism where spatial parallelism is achieved by an efficient
parallel sweep, angular efficiency achieved by pipe-lining, and energy
parallelism achieved by decoupling the groups. With the $SP_N$
equations as a full matrix system, we now can parallelize the problem
as prescribed by the linear solver, which may be significantly more
scalable than current $S_N$ transport practices.

\subsection{Solutions for the Navier-Stokes Equations}
\label{subsec:ns_motiviation}

### 2013-06-11 doc/dissertation/document/chapters/mc_background/mc_background.tex
adding some example figures for domain decomposed monte carlo

Choosing the infinity norm of the operator as defined in
Eq~(\ref{eq:matrix_infinity_norm}), the row normalized probability
matrix will yield a norm of 1 giving the following inequality for
relating $\ve{G}$ and $\ve{H}$:

>>>>

Choosing the infinity norm of the operator, the row normalized
probability matrix will yield a norm of 1 giving the following
inequality for relating $\ve{G}$ and $\ve{H}$:

### 2013-06-11 doc/dissertation/document/chapters/nonlinear_problem/nonlinear_problem.tex
restructuring verification results
adding some example figures for domain decomposed monte carlo

\chapter{Monte Carlo Solution Methods for the Navier-Stokes Equations}
\label{ch:nonlinear_problem}
Nonlinear transport problems are a common occurrence in single and
multiple physics problems. Systems of partial differential equations
such as those that describe fluid flow or more general transport
processes when discretized by conventional methods yield discrete sets
of stiff equations with nonlinearities present in the
variables. Traditionally, such systems have been solved by linearizing
them in a form where the nonlinearities in the variables are
eliminated and more traditional linear methods can be used for
solutions. Often characterized as segregated methods where physics
operators are split and their action on the system approximated in
steps, such methods lack consistency and accuracy in resolving the
nonlinear component of the solution. In the last 30 years, fully
consistent nonlinear methods based on Newton's method have become more
popular and many advances have been made in the computational physics
field to employ these methods.

@@

\subsection{Parallel FANM Implementation}
\label{subsec:parallel_fanm}
Like the parallelization of the MCSA method described in
\S~\ref{sec:parallel_stochastic_methods}, a parallel FANM method
relies on a basic set of parallel matrix-vector operations as well as
the global residual and Jacobian assembly procedure described in
\S~\ref{subsubsec:automatic_differentiation}. Consider the Newton
iteration scheme in Eq~(\ref{eq:newton_iteration}). We must first
assemble the linear system in parallel through the element-wise
function evaluations to generate both the global Jacobian operator and
the global residual vector on the right hand side. Per Bartlett's
work, efficient and automated parallel mechanisms are available to do
this through a sequence of scatter/gather operations. With these tools
available for residual and Jacobian generation, the remainder of the
parallel procedure is simple, with the linear Newton correction system
solved using the parallel MCSA method as previously described and the
Newton correction applied to the previous iterate's solution through a
parallel vector update operation.

>>>>

\chapter{Monte Carlo Synthetic Acceleration Methods for the Navier-Stokes Equations}
\label{ch:nonlinear_problem}
Nonlinear transport problems are a common occurrence in single and
multiple physics problems. Systems of partial differential equations
such as those that describe fluid flow or more general transport
processes when discretized by conventional methods yield discrete sets
of stiff equations with nonlinearities present in the
variables. Traditionally, such systems have been solved by linearizing
them in a form where the nonlinearities in the variables are
eliminated and more traditional linear methods can be used for
solutions. Often characterized as segregated methods where physics
operators are split and their action on the system approximated in
steps, such methods lack consistency and accuracy in resolving the
nonlinear component of the solution. In the last 30 years, fully
consistent nonlinear methods based on Newton's method have become more
popular and many advances have been made in the computational physics
field to employ these methods.

@@

\subsection{Parallel FANM Implementation}
\label{subsec:parallel_fanm}
A parallel FANM method relies on a basic set of parallel matrix-vector
operations as well as the global residual and Jacobian assembly
procedure described in
\S~\ref{subsubsec:automatic_differentiation}. Consider the Newton
iteration scheme in Eq~(\ref{eq:newton_iteration}). We must first
assemble the linear system in parallel through the element-wise
function evaluations to generate both the global Jacobian operator and
the global residual vector on the right hand side. Per Bartlett's
work, efficient and automated parallel mechanisms are available to do
this through a sequence of scatter/gather operations. With these tools
available for residual and Jacobian generation, the remainder of the
parallel procedure is simple, with the linear Newton correction system
solved using the parallel MCSA method as previously described and the
Newton correction applied to the previous iterate's solution through a
parallel vector update operation.
